---
title: "pvcm"
author: "Joonho Gong"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pvcm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pvcm)
library(spls)
```

## Functions

### pvcm
`pvcm` is a function to fit a principal varying coefficient model with adaptive weights. Parallel computting for multi-core CPU is built-in to reduce computational time. 

#### Usage
```{R}
set.seed(12345)
data(yeast)

## Preparation for design matrix X and the long y
# 1. y_vec
y_ready <- c( t( yeast$y ) )
n <- length(y_ready)

# 2. X_mat
p_org <- dim(yeast$x)[2]
X_temp <- apply( scale(yeast$x), 1, FUN = function(x){
  matrix( rep(x, 18), 18, p_org, byrow=TRUE)
}, simplify = FALSE)

X_mat <- matrix(NA, n ,p_org)
for(i in 1:length(X_temp)){
  X_mat[ ( 1 + 18 * (i- 1) ):(18 * i) , ] <- X_temp[[i]]
}

X_ready <- cbind( rep(1, n), X_mat )
colnames(X_ready) <- c( "intercept", colnames(yeast$x) )

## Fit PVCM
pvcm_fit <- pvcm(y_sample = y_ready, X_sample = X_ready, 
                 U_sample = rep(seq(0,1,length.out = 18), dim(yeast$x)[1]), 
                 knots_val = 7, 
                 control = list(num_tau_1=20, num_tau_2=20, max.iter=100, tol=1e-5)
                 )

true_TFs <- c("intercept",
              "ACE2_YPD", "SWI4_YPD", "SWI5_YPD", "SWI6_YPD", "MBP1_YPD", "STB1_YPD", "FKH1_YPD",
              "FKH2_YPD", "NDD1_YPD", "MCM1_YPD", "ABF1_YPD", "BAS1_YPD", "CBF1_YPD", "GCN4_YPD",
              "GCR1_YPD", "GCR2_YPD", "LEU3_YPD", "MET31_YPD", "REB1_YPD", "SKN7_YPD", "STE12_YPD")
# Experimentally confirmed covariates
sort(true_TFs)

true_positive <-  sum(pvcm_fit$selected_covariates %in% true_TFs) - 1
# Not consider the intercept

true_negative <- p_org - (length(true_TFs) - 1) - sum(!(pvcm_fit$selected_covariates %in% true_TFs))
# except intercept

accuracy <- (true_positive + true_negative) / p_org
# 0.754717
```
#### Argument
**y_sample:**  A numeric vector of response variable.

**X_sample:**  A design matrix of covariates including an intercept if it is included in your model.

**U_sample:**  A numeric vector of index variable.

**num_cores:**  The number of CPU cores to implement parallel computing. Default NULL will use the total number of cores - 1).

**knots_val:**  A numeric value for the knots argument of select.knots function in the face R package. Default is 7 for cubic B-spline. 

**control:**  A list that contains two numbers of tuning parameters except 0 and the stopping criteria for DYS algorithm. There are 4 elements: **num_tau_1**, **num_tau_2**, **max.iter**, and **tol**. First two are the number of candidates for each penalty except 0. Default is 20 and zero will be automatically added. The total number of iterations, **max.iter**, and numeric tolerance, **tol**, are stopping criteria of DYS algorithm with default values 100 and 1e-5, respectively. 

#### Value
**Theta_sparse:**  The estimated Theta matrix with zero rows. 

**Theta_low_rank:**  The estimated Theta matrix of low rank. 

**Theta:**  The estimated Theta matrix from DYS algorithm close to **Theta_sparse:** and **Theta_low_rank**.

**rank:**  The estimated rank of Theta matrix.  

**U:**  The matrix of left singular vectors of **Theta_low_rank**.

**V:**  The matrix of right singular vectors of **Theta_low_rank**.

**D:**  The diagonal matrix of singular values of **Theta_low_rank**.

**B_U_mat:**  A matrix of orthogonal basis functions evaluated at **U_sample**. 

**fitted_values:**  A vector of fitted values corresponding to **y_sample**. 

**selected_covariates:**  The names of selected covariates by pvcm. 


### dys_pvcm
`dys_pvcm` is a function specifically developed for the principal varying coefficient models.


#### Argument
**y_vec:**  A numeric vector of response variable.

**M_mat:**  A matrix whose row is the kroneker product between the covariates X_i and the evaluated basis functions at U_i.

**tau:**  A vector of the two tuning parameters.

**Theta_initial:**  An initial value for the Theta matrix.

**weights_ann:**  A vector of weights for the adaptive nuclear norm penalty. 

**weights_agl:**  A vector of weights for the adaptive group lasso penalty. 

**gamma_val:**  A numeric value for one over Lipschitz constant.

**max.iter:**  A maximum iteration number.

**tol:**  A numeric value for tolerance of relative deviation in order to stop DYS iteration.



#### Value
**Theta_sparse:**  The estimated Theta matrix with zero rows. 

**Theta_low_rank:**  The estimated Theta matrix of low rank. 

**Theta:**  The estimated Theta matrix from DYS algorithm close to **Theta_sparse:** and **Theta_low_rank**.

**rank:**  The estimated rank of Theta matrix.  

**U:**  The matrix of left singular vectors of **Theta_low_rank**.

**V:**  The matrix of right singular vectors of **Theta_low_rank**.

**D:**  The diagonal matrix of singular values of **Theta_low_rank**.

**converge:**  A logical value to indicate convergence. 

**rank_trace:**  A vector of estimated rank of Theta matrix over iteration. 

**iteration:**  The total number of iterations

**rel_diff:**  A vector of relative differences over iteration. 

### plot_pvcm
`plot_pvcm` is a function to display estimated varying coefficient functions and pricipal functions. 

#### Usage
```{R}

par(mar = c(3.8, 2, 1.8, 1.8) )
## Varying coefficient functions
plot_pvcm(pvcm_fit, pallet_size = c(4,4), X_sample = X_ready,
          U_sample = rep(seq(0,1,length.out = 18), dim(yeast$x)[1]),
          beta_plot = TRUE, true_covariates = true_TFs
          )

## Principal functions
plot_pvcm(pvcm_fit, pallet_size = c(2,2), 
          U_sample = rep(seq(0,1,length.out = 18), dim(yeast$x)[1]),
          beta_plot = FALSE
          )
```


#### Argument
**pvcm_fit:**  A list output of pvcm function.

**pallet_size:**  A vector of two positive integer elements for the number of rows and columns of pallet.

**X_sample:**  A design matrix of covariates with an intercept if it is included in your model.

**U_sample:**  A numeric vector of index variable.

**beta_plot:**  A logical value to indicate whether the coefficient functions (TRUE) or the principal functions (FALSE) are displayed. 

**true_covariates:** A vector of names for the known true covariates. Default NULL will display the plots for all selected covariates when **beta_plot** is TRUE. 



